---
title: "classifiers_R"
author: 
  name: Manuel Schreiber
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    df_print: paged
  html_notebook: default
  latex_engine: pdflatex
  pdf_document: default
  header-includes:
   - \usepackage{amsmath}
   - \usepackage{amsfonts}
---
```{r}
library(tidyverse) # collection of packages for data analysis including: ggplot2(data visualization package), dplyr (data manipulation package), tidyr, readr, purrr, tibble, stringr, forcats

library(graphics) # base graphics (e.g. histogram)

library(stats) # computing distribution functions, distribution quantiles; random number generation

library(e1071) # package for the Naive Bayes Classifier

library(yardstick) # package including the balanced accuracy score

library(caret) # classification and regression training package

library(randomForest) # Random Forest ensemble method

library(xgboost) # xgboosting ensemble method

```



## changing the WD

```{r}
setwd("/Users/Manu/Desktop/TUM_Master_Mgt_Technology/TUM_WS_20/Business Analytics/Analytics Cup")

```

```{r}
# reading the training and test data sets into R
train = read_csv("train_physicians.csv")
test = read_csv("test_physicians.csv")

```

```{r}
str(train)
```

```{r}
# subsetting the training set for model training and model evaluation

# labels 
y_train = train$ownership_interest[0:4000]

y_test = train$ownership_interest[4001:length(train$ownership_interest)]

str(y_train);

str(y_test)


# features 

# subsetting columns from the train dataframe                 
X_train = subset(train, select=c(total_payments, number_of_payments))   

X_train <- X_train[0:4000,] # subsetting the first 4000 rows
                 

X_test = subset(train, select=c(total_payments, number_of_payments)) 
X_test <- X_test[4001:4788,]

str(X_train);

str(X_test)

```

```{r}
# subsetting the entire training data for the NB model

train_subset = as.data.frame(cbind(y_train,X_train))

names(train_subset)[1] <- "ownership"

str(train_subset);

test_subset = as.data.frame(cbind(y_test,X_test))

names(test_subset)[1] <- "ownership"

str(test_subset)

```



```{r}
# creating a naive Bayes prediction model
# arguments of naiveBayes():
# data: dataframe of predictors
##nb <- naiveBayes(y = y_train, x = X_train, data = X_train, laplace=1)

nb <- naiveBayes(formula = ownership ~ total_payments + number_of_payments, data= train_subset, laplace = 1)

nb
```

TO DO: ONLY NA PREDICTIONS

```{r}
# predictions for the test dataset
##preds_nb = predict(nb, newdata= test_subset, type="class")[0:length(y_test)]

##preds_nb



```



```{r}
# confusion matrix
##table(true=y_test,prediction=predictions)
```


*Logistic regression*

To perform logistic regression, we need to code the response variables into integers. This can be done using the factor() function

Useful link:
http://rstudio-pubs-static.s3.amazonaws.com/74431_8cbd662559f6451f9cd411545f28107f.html



```{r}
## fitting a Generalized Linear Model (GLM) (Logistic Regression) 
log_reg = glm(y_train ~ total_payments + number_of_payments, data=X_train, family=binomial(link="logit"))

summary(log_reg)
```

```{r}
dim(X_test);
dim(X_train)
```


```{r}
# predicted probablities on the test set
preds_proba_lr <- predict(log_reg,newdata=X_test,type='response')[0:length(y_test)]

head(preds_proba_lr);


# predicted classes using a cutoff value of p = 0.5
preds_lr <- rep(0, length(y_test)) # initializing a preds_lr object with 0s

preds_lr[preds_proba_lr > .5] = 1;

length(preds_lr)

table(preds_lr) # frequency table of predictions


```



```{r}
# Calculating a cross-tabulation of observed and predicted classes with associated statistics
accuracy(data = X_test, truth = factor(y_test), estimate = factor(preds_lr))

```


```{r}
# confusion matrix
table(true=y_test,prediction=preds_lr)
```


```{r}
# computing the balanced accuracy score
bal_accuracy(data = X_test, truth = factor(y_test), estimate = factor(preds_lr))

```


Random Forest


Random Forest (RF) models improve on the idea of bagging by de-correlating the trees

RF models build a number of regression trees/classification trees on bootstrapped training samples, but when building these trees, each time a split in a tree is considered, only a random selection of m predictors (not all) are allowed as split candidates from the full set of p predictors (usually m=p)

The key idea here is that by randomizing the set of allowed predictors, the resulting trees are less correlated with each other.

```{r}
# setting the seed for reproducibility
set.seed(123)
```


```{r}
# fitting a random forest model
rf = randomForest(factor(y_train) ~total_payments + number_of_payments, data=X_train, ntree=100, proximity=T)

rf


```

```{r}
# feature importance

importance(rf)

```

```{r}
dim(X_test)
```


```{r}
preds_rf = predict(rf, newdata=X_test, type='response')[0:length(y_test)]

head(preds_rf)

```

```{r}
# confusion matrix
table(true=y_test,prediction=preds_rf)
```


```{r}
# Calculating a cross-tabulation of observed and predicted classes with associated statistics
accuracy(data = X_test, truth = factor(y_test), estimate = factor(preds_rf))

# not that useful as a metric for imbalanced classification data!
```



```{r}
# computing the balanced accuracy score
bal_accuracy(data = X_test, truth = factor(y_test), estimate = factor(preds_rf))

```


*XGBoosting (extreme gradient boosting) model*

https://xgboost.readthedocs.io/en/latest/R-package/xgboostPresentation.html

Combination of regularization, boosting and a Random Forest

XGB supports various objective functions, including regression, classification and ranking

Input Type: it takes several types of input data:
(i) Dense Matrix: R’s dense matrix, i.e. matrix ;
(ii) xgb.DMatrix: its own class (recommended)
XGBoost offers a way to group them in a xgb.DMatrix. You can even add other meta data in it. This will be useful for the most advanced features we will discover later.

Customization: it supports customized objective functions and evaluation functions.


```{r}
# we have already set the seed further up in line 209

# formatting the training data

train_features <- as.matrix(train_subset[c("total_payments","number_of_payments")])

class(train_features);

train_label <- train_subset$ownership

# formatting the test data

test_features <- as.matrix(train_subset[c("total_payments","number_of_payments")])

test_label <- test_subset$ownership


       
```

```{r}
# inspecting the training and evaluation data

table(train_label)

# majority and minority class shares
c_major <- sum(train_label == 0)/(sum(train_label == 0) + sum(train_label == 1))

c_minor <- sum(train_label == 1)/(sum(train_label == 0) + sum(train_label == 1))

c_major;
c_minor





```


```{r}
## Gridsearch code for R (caret package) ##
 
# step 1: setting up a grid of hyperparameters

# creating a data frame from all combinations of the supplied vectors or factors
# using the expand.grid() function
xgb_tune_grid = expand.grid(
nrounds = c(10,100), # no. of passes of the data, i.e. no. of boosting iterations
eta = c(0.01,0.1,0.3), # learning rate
max_depth = c(5,10), # max no. of leaves for each tree
gamma = c(1,2), # min loss reduction required to make a further leaf node partition
colsample_bytree = c(0.7,0.9), # subsample of columns for a single tree
min_child_weight = c(5,10), # min sum of instance weight needed in a child node
subsample = c(0.7,0.8) # subsample of rows considered (RF component),
)

# step 2: setting the training control parameters (type of CV, # of folds, etc)
 
## trainControl() sets the training control parameters for the train() function:##

xgb_trcontrol_params = trainControl(
method = "cv", # cross validation
number = 3, # number of folds
verboseIter = TRUE,
returnData = FALSE,
returnResamp = "all",                                                        
# save losses across all models
#classProbs = FALSE,                                                           
allowParallel = TRUE # set to FALSE for reproducibility
)

# step 3: setting the input nad the method for the models to be fitted during the GSCV
 
## train the model for each parameter combination in the grid and use a CV to evaluate##

gscv_fit = caret::train(
x = train_features,
y = factor(train_label), # using a factor as label for classification
trControl = xgb_trcontrol_params,
tuneGrid = xgb_tune_grid,
method = 'xgbTree' # extreme gradient boosted trees
)


```

Jan 19: Selecting tuning parameters
Fitting nrounds = 10, max_depth = 5, eta = 0.01, gamma = 1, colsample_bytree = 0.7, min_child_weight = 5, subsample = 0.7 on full training set

```{r}
# best gscv score
gscv_fit$results
```


```{r}
# fitting the optimal xgb model (after the GSCV)

xgb_opt <- xgboost(data = train_features, 
              label = train_label,
              objective="binary:logistic", # has to be changed to balanced accuracy
              eta = 0.01,
              gamma = 1,
              max_depth = 5,
              min_child_weight = 5,
              subsample = 0.7,
              colsample_bytree = 0.7,
              nrounds=10,
              verbose = 2)

```

Baseline XGB boosting classifier

```{r}
# setting up the xgboost model (baseline xgb model)
  
xgb_base <- xgboost(data = train_features, 
              label = train_label,
              objective="binary:logistic",
              learning_rate = 0.5,
              lambda = 1.95,
              max_depth = 7,
              subsample = 0.7,
              colsample_bytree = 0.7,
              colsample_bylevel = 0.7,
              nround=20,
              verbose = 1)

```


```{r}
# feature importance plot
xgb_importance <- xgb.importance(model = xgb_base, label = train_label)

xgb_importance # data table

# plot of top 10 most important features
xgb.plot.importance(importance_matrix = xgb_importance, top_n = 10) 
```


```{r}
# prediction on the test set
preds_proba_xgb <- predict(xgb_base, test_features)[0:length(y_test)] 

head(preds_proba_xgb);


# predicted classes using a cutoff value of p = 0.5
preds_xgb <- rep(0, length(y_test)) # initializing a preds_xgb object with 0s

preds_xgb[preds_proba_xgb > .5] = 1;

length(preds_xgb)

table(preds_xgb) # frequency table of predictions

```


```{r}
# confusion matrix
table(true=y_test,prediction=preds_xgb)
```


```{r}
# computing the balanced accuracy score
bal_accuracy(data = X_test, truth = factor(y_test), estimate = factor(preds_xgb))

```




