---
title: "eda_analytics_cup"
author: 
  name: Manuel Schreiber
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    df_print: paged
  html_notebook: default
  latex_engine: pdflatex
  pdf_document: default
  header-includes:
   - \usepackage{amsmath}
   - \usepackage{amsfonts}
---

**UPDATES**

**January 11: payments.csv file from training data updated**

**January 20: hints from coaching session:**
split 5000 data points into train and evaluation set (e.g. 4000/1000)
-> then train the model again using all 5000 data points before making the
predition on the test features belonging to the hidden target variables

Plot categorical features and computing a correlation matrix for numerical variables to detect correlations/find useful features 

**Task**

Predict Ownership Interests of physicians in pharmaceutical companies from regulatory transaction data.

**Task description:**

You are given 1 400 000 instances of transactions which have been received by 6000 physicians in the years 2013 to 2019 (transactions.csv), as well as additional data about these doctors (physicians.csv) and companies (companies.csv). For 5000 of these doctors, transactions related to ownership interest (if any) are included, for the other 1000 (the test set), these kinds of transactions have been removed from the transaction data. For each of these 1000 physicians, your model must make a prediction about whether they have had such ownership interests (prediction=1) or not (prediction=0).

```{r}
# clear workspace
##rm(list=ls())
```




```{r}
# Loading Packages

library(tidyverse) # collection of packages for data analysis including: ggplot2 (data visualization package), dplyr (data manipulation package), tidyr, readr (faster read.csv() functions that create tibble objects instead of data.frames), purrr, tibble (wrapper around data.frames for better printing), stringr, forcats

library(lubridate) # tools to parse and manipulate dates

library(GGally) # ggplot2 extension (for a heat map plot)

library(graphics) # base graphics (e.g. histogram)

library(stats) # computing distribution functions, distribution quantiles; random number generation


library(tidymodels) # collection of packages for modeling and machine learning

library(googledrive) # interface to googledrive

library(descr) # crosstab function for EDA

library(mltools) # one hot encoding function

library(data.table)


```

```{r}
# obtaining your current working directory
getwd()
```


```{r}
# finds all CSV files on your google drive
##drive_find(type = "csv")  

# downloading a file from Gdrive to your working directory
##drive_download("water_qual.csv", overwrite = TRUE)

# uploading a file to Gdrive 
##drive_upload()

# updating a file on Gdrive
##drive_update()

# deleting files on Gdrive
##drive_rm()


```

```{r}
##drive_download("physicians.csv")
#drive_download("payments.csv", overwrite = TRUE)

```

# Exploratory Data Analysis

## Physicians dataframe

```{r}
physicians_df <- read_delim("physicians.csv", delim = ",")
```

```{r}
str(physicians_df)
```


```{r}

head(physicians_df)

# interesting features: id -> way to join with payments df
```



```{r}
# finding the number of NA values by column in the physicians df
physicians_df %>% mutate_all(is.na) %>% summarize_all(sum)
```

```{r}
# all colnames  
colnames(physicians_df) 

# reduced dataframe -: removing variables
physicians_df_red <- subset(physicians_df, select = -c(First_Name,Middle_Name,Last_Name, Name_Suffix))
                      

# # storing all physician id's for the training set
# id_train <- physicians_df_red$id[physicians_df_red$set=="train"]
# 
# # storing all physician id's for the test set
# id_test <- physicians_df_red$id[physicians_df_red$set=="test"]
# 
#                         
  


```


*Plotting categorical features and computing a correlation matrix for numerical variables*

```{r}
#crosstab(dep = physicians_df$id, indep = physicians_df$Primary_Specialty)
```


## payments dataframe

```{r}
payments_df <- read_delim("payments.csv", delim = ",")
```

```{r}
str(payments_df)
# interesting features. Physician_ID, Company_ID, Ownership_Indicator
```

*Plotting categorical features and computing a correlation matrix for numerical variables*

```{r}
# correlation matrix of numerical features
cor(select(payments_df, where(is.numeric)))
```
correlations worth looking at: Company_ID - Record_Id

```{r}
# barplot of the frequency table of the categorical feature "Country of Travel"
table(payments_df$Ownership_Indicator, payments_df$Country_of_Travel)

# plotting the top 5 most frequent values of the categorical feature Country of Travel
barplot(sort(table(payments_df$Country_of_Travel), decreasing = TRUE)[1:5], cex.names = 0.8)

```


```{r}
# cross tab plot of ownership interest and certain features with legend
barplot(with(payments_df, table(Ownership_Indicator, Third_Party_Recipient, exclude = "NA")), beside = TRUE, legend.text = TRUE, main = "Crosstab", cex.names = 0.8)
```

```{r}
# cross tab plot of ownership interest and certain features with legend
barplot(with(payments_df, table(Ownership_Indicator, Charity)), beside = TRUE, legend.text = TRUE, main = "Crosstab", cex.names = 0.8)

# cross tab plot (way 2)
crosstab(dep=payments_df$Ownership_Indicator, indep= payments_df$Charity)

# cross table
with(payments_df, table(Ownership_Indicator, Charity))

```



```{r}
# Cross table of the values of the categorical feature State of Travel and ownership
table(payments_df$Ownership_Indicator, payments_df$State_of_Travel)

# plotting the top 10 most frequent values of the categorical feature State of Travel
barplot(sort(table(payments_df$State_of_Travel), decreasing = TRUE)[1:10])


# cross tab plot of ownership interest and certain features with legend
# horiz: if T, bars are drawn horizontally, if F, bars are drawn vertically
barplot(with(payments_df, table(Ownership_Indicator, State_of_Travel)), beside = TRUE, legend.text = TRUE, main = "Crosstab", cex.names = 0.6, horiz=F)

```

```{r}
# cross tab plot of ownership interest and certain features with legend
# cex.names: expansion factor for the axis labels
barplot(with(payments_df, table(Ownership_Indicator, Related_Product_Indicator, exclude = "NA")), beside = TRUE, legend.text = TRUE, main = "Crosstab", cex.names = 0.8)

print(unique(payments_df$Related_Product_Indicator))

# Cross table of the values of the categorical feature Realted_Product_Indicator and ownership
table(payments_df$Ownership_Indicator, payments_df$Related_Product_Indicator)


```


One-hot encoding is the process of converting a categorical variable with multiple categories into multiple variables, each with a value of 1 or 0.

```{r}
# Cross table of the values of the categorical feature Transfer of Value and ownership
table(payments_df$Ownership_Indicator, payments_df$Form_of_Payment_or_Transfer_of_Value)


unique(payments_df$Form_of_Payment_or_Transfer_of_Value)


# cross tab plot of ownership interest and certain features with legend
# cex.names: expansion factor for the axis labels
barplot(with(payments_df, table(Ownership_Indicator, Form_of_Payment_or_Transfer_of_Value)), beside = TRUE, legend.text = TRUE, main = "Crosstab", cex.names = 0.6)


# one hot encoding of the 7 categories

##oh_df <- payments_df$Physician_ID

##oh_df$form_of_payment <- factor(payments_df$Form_of_Payment_or_Transfer_of_Value)

##new_data <- one_hot(as.data.table(oh_df))

##str(new_data)

```

```{r}
# unique values (14 in this case)
unique(payments_df$Nature_of_Payment_or_Transfer_of_Value)

# cross table
table(payments_df$Ownership_Indicator,payments_df$Nature_of_Payment_or_Transfer_of_Value)

# cross tab plot of ownership interest and certain features with legend
# cex.names: expansion factor for the axis labels
barplot(with(payments_df, table(Ownership_Indicator, Nature_of_Payment_or_Transfer_of_Value)), beside = TRUE, legend.text = TRUE, main = "Crosstab", cex.names = 0.4)

# plotting the top 5 most frequent values of the categorical feature Nature of Payment
barplot(sort(table(payments_df$Nature_of_Payment_or_Transfer_of_Value), decreasing = TRUE)[1:5], cex.names = 0.6)


```


```{r}
# cross tab plot of ownership interest and certain features with legend
# cex.names: expansion factor for the axis labels
barplot(with(payments_df, table(Ownership_Indicator, Third_Party_Covered)), beside = TRUE, legend.text = TRUE, main = "Crosstab", cex.names = 0.6)
```



```{r}
# example one hot ecoding
set.seed(555)
data <- data.frame(
  Outcome = seq(1,100,by=1),
  Variable = sample(c("Red","Green","Blue"), 100, replace = TRUE)
)


newdata <- one_hot(as.data.table(data))

newdata
```


```{r}
# PCA attempt

trainset <- select(payments_df, where(is.numeric))

str(trainset)


# PCA
pca_trainset = trainset 
#pca_testset = testset
pca = prcomp( pca_trainset, scale = T )

# variance
pr_var = ( pca$sdev )^2 

# % of variance
prop_varex = pr_var / sum( pr_var )

# Plot (proportion of variance explained by principal component)
plot( prop_varex, xlab = "Principal Component", 
                  ylab = "Proportion of Variance Explained", type = "b" )


# cummulative variance explained plot
plot( cumsum( prop_varex ), xlab = "Principal Component", 
                            ylab = "Cumulative Proportion of Variance Explained", type = "b" )
```



```{r}
unique(payments_df$Ownership_Indicator)

unique(payments_df$Nature_of_Payment_or_Transfer_of_Value)

# for loop which returns all unique values of each feature in the payments df
for (i in (1:length(payments_df))) { 
 print(unique(payments_df[i]))
}

```



```{r}
# finding the number of NA values by column in the payments df
payments_df %>% mutate_all(is.na) %>% summarize_all(sum)
```

```{r}
summary(payments_df)
```

## companies dataframe

```{r}
companies_df <- read_delim("companies.csv", delim = ",")
```

```{r}
str(companies_df)
```

```{r}
# renaming columns in the company df; rename(df_name, new_col_name = old_col_name)
companies_df <- rename(companies_df, c(Company_name = Name, Company_state = State))

str(companies_df)
```



```{r}
# uploading the markdown file to gdrive
##drive_upload("eda.html")
```

*Merging 2 dataframes by a common column*

```{r}
# renaming the id column in the physicians_df
# renaming a column way 1 using a base R function
##names(physicians_df)["id"] <- "Physician_ID"

# renaming a column way 2 using the dplyr function rename() and updating the dataframe
physicians_df <- rename(physicians_df,  Physician_ID = id)

colnames(physicians_df)

# joining the new payments_df and companies_df by Company_ID
df <- merge(payments_df, companies_df, by="Company_ID")

# joining physicians_df and payments_df by Physician_ID using merge()
df_full <- merge(physicians_df, df, by="Physician_ID", all=T)

length(df_full$Ownership_Indicator)
length(df_full$Company_name)

```

```{r}
#str(df_full)



#barplot(table(df_full$Ownership_Indicator, df_full$Company_name))

# cross tab plot of ownership interest and certain features with legend
# cex.names: expansion factor for the axis labels
barplot(with(df_full, table(Ownership_Indicator, Company_name)), beside = TRUE, legend.text = TRUE, main = "Crosstab", cex.names = 0.6)



```


